### 2023-01-24 SI.tex
supporting information tex file (#30)

% document
\documentclass[11pt, full]{article}
\usepackage[letterpaper, portrait, margin=0.75in]{geometry}
\usepackage{setspace}
\usepackage{color}

% text
\usepackage[utf8]{inputenc}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}
\usepackage{enumitem}
\renewcommand{\familydefault}{\sfdefault}
\newcommand{\RomanNumeral}[1]{\textrm{\uppercase\expandafter{\romannumeral #1\relax}}}

% math
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage[cm]{sfmath}
\usepackage{commath}
\usepackage{multirow}
\DeclareMathAlphabet{\mathpzc}{OT1}{pzc}{m}{it}

% graphics
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{epstopdf}
\usepackage{xpatch}
\graphicspath{{./figures/}}

% "S" prefix
\renewcommand{\theequation}{S\arabic{equation}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thetable}{S\arabic{table}}

% each section begins new page
\let\stdsection\section
\renewcommand\section{\clearpage\stdsection}

% force all floats to center (see https://tex.stackexchange.com/a/53383)
\makeatletter
\g@addto@macro\@floatboxreset{\centering}
\makeatother

% bibliography
\usepackage[backend=bibtex, natbib=true, url=false, sorting=none, maxbibnames=99]{biblatex}
\bibliography{references}

% hyperref
\usepackage[colorlinks=true, linkcolor=black, urlcolor=blue, citecolor=black, anchorcolor=black]
  {hyperref}
\usepackage[all]{hypcap}  % helps hyperref work properly

\begin{document}
\pagenumbering{gobble}

\begin{center}
  \LARGE

  \textit{Supporting Information} \\
  The yaq Project: \\
  Standardized Software Enabling Flexible Instrumentation

  \normalsize

  \textit{Kyle F. Sunden, Daniel D. Kohler, Kent A. Meyer, Peter L. Cruz Parrilla, \\
    John C. Wright, and Blaise J. Thompson*}

  Department of Chemistry, University of Wisconsin--Madison\\
  1101 University Ave., Madison, Wisconsin 53706
\end{center}

\vspace{\fill}

*Corresponding Author \\
\hspace*{2ex} email: blaise.thompson@wisc.edu \\
\hspace*{2ex} phone: (608) 263-2573

\pagebreak
\setcounter{page}{1}
\pagenumbering{arabic}
\renewcommand{\thepage}{S\arabic{page}}

\pagebreak
\renewcommand{\baselinestretch}{0.75}\normalsize
\tableofcontents
\renewcommand{\baselinestretch}{1.0}\normalsize

\section{Package size analysis}

\clearpage

\printbibliography

\end{document}